---
title: "Support Vector Machines"
author: "Nick Phynn"
date: "12/20/2024"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = FALSE}
library(caret)
```


```{r include=FALSE}
data <- read.csv("airquality.csv")
str(data)
sum(is.na(data))
unique(data$RH)

library(tidyr)
data1 <- 
  data %>%
  drop_na()
sum(is.na(data1))
str(data1)
```


## A sample of 100 observation from the air quality data set will be used train and predict Support Vector Machines for the best method of accuracy: Linear, Non-Linear, Radial, Poly.

```{r}
data1$class <- ifelse(data1$RH < mean(data1$RH), 'No', 'Yes')
data1$class <- factor(data1$class, levels = c('No','Yes'), labels = c("No","Yes"))
table1 <- table(data1$class)
realdata <- data1[1:100, ]
str(realdata)
```



Split the data into training and testing resource

```{r}
index <- createDataPartition(realdata$RH, p =0.70, list = FALSE)
training <- realdata[index,]
valid <- realdata[-index,]
```



Control harness for prediction

```{r}
control <- trainControl(method="cv", number=10)
metric <- "Accuracy"
```

# LINEAR SVM



Model Training: Linear SVM model using the binomial family funciton with 72 samples, 10 predictor and 2 classes. The accuracy is 97% and Kappa is 94%

```{r}
set.seed(7)
fit.svm <- train(class~., data = training, method = "svmLinear", family=binomial(), trControl=control, metric=metric)
fit.svm
```




Model Prediction: The model has an accuracy of 100% and Kappa of 100%. The confusion matrix confirming no false positive or negatives

```{r}
data.pred <- predict(fit.svm, newdata = valid)
cm <- confusionMatrix(as.factor(data.pred), reference = as.factor(valid$class), mode = "prec_recall")
print(cm)
```




# NON-LINEAR SVM


Model Training: Non-Linear SVM model with 72 samples, 10 predictor and 2 classes. The accuracy is 97% and Kappa is 94%



```{r}
set.seed(7)
fit.LR <- train(class~., data = training, method = "svmLinear", trControl=control, metric=metric)
fit.LR
```




Model Prediction: The model has an accuracy of 100% and Kappa of 100%. The confusion matrix confirming no false positive or negatives



```{r}
data.pred1 <- predict(fit.LR, newdata = valid)
cm <- confusionMatrix(as.factor(data.pred1), reference = as.factor(valid$class), mode = "prec_recall")
print(cm)
```




# RADIAL SVM



Model Training: Radial SVM model with 72 samples, 10 predictor and 2 classes. The accuracy is 97% and Kappa is 93% with C = 1 the optimal model out of the three predictions



```{r}
set.seed(7)
fit.LogitBoost <- train(class~., data = training, method="svmRadial", metric=metric, trControl=control)
fit.LogitBoost 
```




Model Prediction: The model has an accuracy of 96% and Kappa of 92%. The confusion matrix displayed 1 false positives which impacted the prediction


```{r}
data.pred2 <- predict(fit.LogitBoost, newdata = valid)
cm <- confusionMatrix(as.factor(data.pred2), reference = as.factor(valid$class), mode = "prec_recall")
print(cm)
```







# POLY SVM

*Model Training: Poly SVM model with 72 samples, 10 predictor and 2 classes. The accuracy is 97% and Kappa is 92% with the optimal model used were degree = 3, scale = 0.1 and C = 0.5.*



```{r}
set.seed(7)
fit.plr <- train(class~., data = training, method="svmPoly", metric=metric, trControl=control)
fit.plr 
```




Model Prediction: The model has an accuracy of 96% and Kappa of 92%. The confusion matrix displayed 1 false negative which impacted the prediction



```{r}
data.pred3 <- predict(fit.plr, newdata = valid)
cm <- confusionMatrix(as.factor(data.pred3), reference = as.factor(valid$class), mode = "prec_recall")
print(cm)
```





# COMPARE SVM MODELS


Linear SVM model had the best mean accuracy (97%) and mean kappa (94%) compared to the other model


```{r}
results <- resamples(list(SVM = fit.svm, LogR = fit.LR, LogitBoost=fit.LogitBoost))
summary(results)
dotplot(results)
```

